Casual Bandits

### 2016-01-15 causal-bandit/icml-2015/introduction.tex
Introduce general causal bandit problem.

One obvious approach would be to cast the above example as a stochastic multi-armed bandit problem~\cite{Robbins1952} -- each of the six possible interventions is an arm and the crop yield is the reward -- and apply an algorithm such as UCB~\cite{Auer1995} with well understood regret guarantees.
However, as we show in \S\ref{sec:results}, this approach ignores the extra information available in the non-intervened variables and subsequently yields a worst-case regret with a sub-optimal dependence on the number of possible interventions.

There has been a significant amount of recent work which proposes alternative modes of feedback within the bandit setting \cite{TODO} as well as results for partial monitoring which, as first glance, may appear applicable. As we show in Section~\ref{}, applying these results to our problem also exhibit a sub-optimal dependence on the number of available interventions.

>>>>

This paper can be seen as a first step towards formalising and solving problems such as these. In \S\ref{sec:defs} we introduce \emph{causal bandit problems} and focus on a

\subsection{Related Work}

One obvious approach would be to cast the above example as a stochastic multi-armed bandit problem~\cite{Robbins1952} -- each of the six possible interventions is an arm and the crop yield is the reward -- and apply an algorithm such as UCB~\cite{Auer1995} with well understood regret guarantees.
However, as we show in \S\ref{sec:results}, this approach ignores the extra information available in the non-intervened variables and subsequently yields a worst-case regret after $T$ round of $\bigomega{\sqrt{NT}}$ with a sub-optimal dependence on $N$, the number of possible interventions.

\todom{Note that feedback work as uses graphs}

There has been a significant amount of recent work which proposes alternative modes of feedback within the bandit setting \cite{TODO} as well as results for partial monitoring which, at first glance, may appear applicable. As we show in Section~\ref{}, applying these results to our problem also exhibit a sub-optimal dependence on the number of available interventions.

### 2016-01-15 causal-bandit/icml-2015/setup.tex
Introduce general causal bandit problem.

\newcommand{\bernoulli}{\operatorname{Bernoulli}}
\newcommand{\dirac}{\operatorname{Dirac}}

@@

We consider two standard performance measures. The first is the cumulative regret, which measures the difference between the reward expected under
the omnipotent strategy that knows the optimal action in advance and the expected reward of the learner. 
\eqn{
\label{eq:regret}
R_T = T \mu^* - \E{\sum_{t=1}^T Y_t}\,.
}
The second performance measure is the simple regret, which measures the performance of the learner on the final round only.
\eqn{
\label{eq:regret-simple}
\simpleregret = \mu^* - \E{\mu_{I_T,J_T}}\,.
}
The simple regret typically makes more sense when the learning agent has a fixed learning budget after which its policy will be fixed indefintely. The cumulative
regret is more appropriate for online agents that may continue to adapt their policy.

>>>>

\newcommand{\bernoulli}{\operatorname{Bernoulli}}
\newcommand{\dirac}{\operatorname{Dirac}}
\renewcommand{\vec}[1]{\boldsymbol{#1}}

We now introduce a novel class of stochastic sequential decision problems which we call \emph{causal bandit problems}. In these problems, rewards are given for repeated interventions on a fixed causal model~\cite{Pearl2000}. 
The \emph{causal model} is specified using a directed acyclic graph over a set of random variables $\mathcal{X}$ that defines a Bayesian network over $\mathcal{X}$ where an edge from variable $X$ to $X'$ is interpreted to mean that a change in the value of $X$ directly causes a change to the value of $X'$. 
An \emph{intervention (of size $n$)}, denoted $do(\vec{X}=\vec{x})$, assigns the values $\vec{x}=\{x_1, \ldots, x_n\}$ to the corresponding variables $\vec{X}=\{X_1, \ldots, X_n\}$ with the empty intervention (where no variable is set) denoted $do()$.
An intervention ``mutilates'' the original graph by removing edges into the variables in $\vec{X}$ and the resulting graph defines a probability distribution $\P{\vec{X}^c | do(\vec{X}=\vec{x})}$ over $\vec{X}^c := \mathcal{X} - \vec{X}$. 
Further details can be found in Chapter 21 of~\cite{Koller2009}.

A learner for a casual bandit problem is given the casual model, a set of \emph{observable} variables $\mathcal{O} \subseteq \mathcal{X}$, a set of \emph{settable} variables $\mathcal{S} \subset \mathcal{X}$, a \emph{reward variable} $Y \in \mathcal{O} \cap \mathcal{S}^c$, and the following game is played over $T$ rounds. In round $t$, the learner selects an intervention $do(\vec{X}_t = \vec{x}_t)$ for $\vec{X}_t \subseteq \mathcal{S}$. Values for all non-intervened variables $\vec{X}^c_t$ are sampled from $\P{\vec{X}^c_t | do(\vec{X}_t = \vec{x}_t)}$ and the learner is shown the sampled values of all the observable variables in $\vec{X}^c_t \cap \mathcal{O}$, and recevies the reward $Y_t$.
We note that classical $K$-armed stochastic bandit problem can be recovered in our framework by considering a causal model with an edge connecting a single settable variable $X$ that can take on $K$ values to a reward variable $Y = r(X) + \eta$ for some arbitrary but unknown, real-valued function $r$ and a noise term $\eta$ that may depend on $X$.

The expected reward for playing $do(\vec{X}_t = \vec{x}_t)$ is $\mu_{\vec{x}_t} := \E{Y_t | do(\vec{X}_t = \vec{x}_t)}$.

The \emph{cumulative regret} measures the difference between the reward expected under
the omnipotent strategy that knows the optimal action in advance and the expected reward of the learner and
is denoted
\eqn{
\label{eq:regret}
R_T = T \mu^* - \E{\sum_{t=1}^T Y_t}\,.
}
The second performance measure is the \emph{simple regret}, which measures the performance of the learner on the final round only.
\eqn{
\label{eq:regret-simple}
\simpleregret = \mu^* - \E{\mu_{I_T,J_T}}\,.
}
The simple regret typically makes more sense when the learning agent has a fixed learning budget after which its policy will be fixed indefintely. The cumulative
regret is more appropriate for online agents that may continue to adapt their policy.

\subsection{Simple Causal Bandit Problem}

In the remainder of this paper we focus on a causal model that is sufficient to demonstrate the separation of regret bounds between algorithms that make use of the additional causal information and those which do not.

@@
